Missing Data: Current Practice in Football Research and Recommendations for Improvement

### 2020-11-05 paper/paper.Rmd
formatting.
mdash edits.

Affiliations

@@

Missing data are values that should have been observed, but were not. Since the
values are unobserved, this can mean the population of interest might not be 
properly sampled, inducing bias into a study's results. Left undetected or poorly handled, 
missing data can undermine the validity and accuracy of research results
[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
missing many ways. For example, accidentally skipping a survey question,
equipment failure, or intentionally not recording values. Before
analysis, missing data must be handled, with values typically deleted or
imputed. There is no universal approach to handling missing data. 
Contextual factors--such as, the study design and objective, and
pattern of missingness--determine how missings should be handled, on a
case-by-case basis [@Sainani2015].

@@

Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`). shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration--especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).

@@

Practice to avoid: *Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. Authors sometimes choose to replace missing 
observations with mean values (e.g., participant or group average) [@Sampson2018; @Cresswell2009]. 
While imputing the mean preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data -- such as linear regression, or K nearest neighbours -- should
be considered.

Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
method used to impute missing data has the potential to
affect the outcome of a study. It is important that authors
understand, and document, how imputation affected the goal of the
analysis (i.e., inference, prediction, or both). For example,
understanding the implications in using mean imputation, compared to
K-nearest neighbour imputation--in terms of the effect on parameter estimates
('significant' versus 'not significant'), and the (un)certainty of the
coefficients (i.e., less/more). When using imputation, authors need to
perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**).

@@

1.  Describe screening procedures (define or example -- think about screened out?).

>>>>

Affiliations:

@@

Missing data are values that should have been observed, but were not. Since the
values are unobserved, this can mean the population of interest might not be 
properly sampled, inducing bias into a study's results. Left undetected or poorly handled, 
missing data can undermine the validity and accuracy of research results
[@Sainani2015; @Nakagawa2008; @Sterne2009]. Data can go
missing many ways. For example, accidentally skipping a survey question,
equipment failure, or intentionally not recording values. Before
analysis, missing data must be handled, with values typically deleted or
imputed. There is no universal approach to handling missing data. 
Contextual factors---such as, the study design and objective, and
pattern of missingness---determine how missings should be handled, on a
case-by-case basis [@Sainani2015].

@@

Of the articles that did not mention missing data `r n_no_mention_missings` / `r n_screened`, `r binom_ci_data_shared$mean_pct`(95% CI; `r binom_ci_data_shared$lower_pct` to `r binom_ci_data_shared$upper_pct`). shared their dataset, making
it difficult to determine whether there was any missingness. While the
absence of a missing data statement does not necessarily mean missing
data are not being reported, it does suggest that this aspect of the
analysis process receives little consideration---especially in contrast
to many of the papers that included (potentially meaningless) tests of
normality (e.g., Kruskall-Wallis test).

@@

Practice to avoid: *Mean imputation.* The default of most statistical software for
handling missing data in modelling is listwise deletion. Listwise
deletion removes entire rows that contain missing observations from the
analysis. At best, listwise deletion reduces the power in analysis by
reducing the sample size. At worst, it introduces bias. For example, if
injured participants are removed from a study on injury prevention.
Imputing values removes the need to conduct listwise deletion. However,
it is critical how these values are imputed, and that the imputation
method is documented. Authors sometimes choose to replace missing 
observations with mean values (e.g., participant or group average) [@Sampson2018; @Cresswell2009]. 
While imputing the mean preserves the study sample size (and point estimates), it also reduces
the variance, can alter the relationship between the variable with
missing observations and other variables, and can bias (underestimate)
standard errors [@Scheffer2002]. Smaller standard errors typically reduce *p*-values, which may lead to
incorrect inference. Mean imputation should be avoided. An imputation
strategy that incorporates information from other related variables in
the data---such as linear regression, or K nearest neighbours---should
be considered.

Practice to avoid: *Not evaluating the effect, or choice, of imputation.* The 
method used to impute missing data has the potential to
affect the outcome of a study. It is important that authors
understand, and document, how imputation affected the goal of the
analysis (i.e., inference, prediction, or both). For example,
understanding the implications in using mean imputation, compared to
K-nearest neighbour imputation---in terms of the effect on parameter estimates
('significant' versus 'not significant'), and the (un)certainty of the
coefficients (i.e., less/more). When using imputation, authors need to
perform a sensitivity analysis (**ref circulation article-Nick, possible reference from dave - https://doi.org/10.1191%2F1740774504cn032oa**).

@@

1.  Describe screening procedures **(define or example---think about screened out?)**.
